---
title: 'Assignment #4: Latin American Trade'
author: "Jina Gil"
date: "4/7/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

Latin America is well-known for its commodity exports, and majority of people easily thinks that Latin America's manufacturing sectors, 
especially high-technology sectors, are comparably weaker than most of nations of the world. 

Is it true? How much the import and export of Latin American manufacturing sectors takes from its GDP? 

## Research Question

- Do Argentina, Brazil, and Chile's main exports are commodity goods, including agriculture and mining?
- How much the manufacturing sectors take part of their GDP per Capita growth? 
- How do their technology-development rate and human capital index look like? 

## Data Selection and Methods

- Trade data is from UN's COMTRADE dataset, at SITC 2 digits. 
- GDP per Capita growth, TFP level (and growth), and human capital index are all driven from PwT 9.1 dataset.

---
## Some pre-works...
```{r}
rm(list=ls())
library(SDSRegressionR)
setwd("C:/Users/Jina Gil/Documents/SNU/Spring 2020/Latin America/Assignments")
library(readxl)
ctr <- read.csv("data/comtrade2.csv", stringsAsFactors = FALSE)
library(dplyr)
ctr %>% arrange(country)
```
---

## First, Argentina
### Exports!

```{r}
arg <- ctr[ctr$country %in% c("Argentina"),]
arg.ex <- arg[arg$tflowcode %in% c(2),]
arg.im <- arg[arg$tflowcode %in% c(1),]
ga.ex <- ggplot(arg.ex, aes(x=Period, y=trade.value, group=comcode, color=comdesc))+geom_line()+theme_bw() + 
  labs(title="Exports", subtitle="Argentina",
       x="Year", y="Values", fill="Types of Exported Goods")
ga.ex
```

## Second, Brazil

```{r}
brz <- ctr[ctr$country %in% c("Brazil"),]
brz.ex <- brz[brz$tflowcode %in% c(2),]
brz.im <- brz[brz$tflowcode %in% c(1),]
gb.ex <- ggplot(brz.ex, aes(x=Period, y=trade.value, group=comcode, color=comdesc))+geom_line()+theme_bw() + 
  labs(title="Exports", subtitle="Brazil",
       x="Year", y="Values", fill="Types of Exported Goods")
gb.ex
```

## Third, Chile
```{r}
chile <- ctr[ctr$country %in% c("Chile"),]
chile.ex <- chile[chile$tflowcode %in% c(2),]
chile.im <- chile[chile$tflowcode %in% c(1),]
gc.ex <- ggplot(chile.ex, aes(x=Period, y=trade.value, group=comcode, color=comdesc))+geom_line()+theme_bw() + 
  labs(title="Exports", subtitle="Chile",
       x="Year", y="Values", fill="Types of Exported Goods")
gc.ex
```

## Export and Import

- Their main exports (except Argentina): **crude materials**. 
- However, all three states also exports lots of **manufactured goods**. 

---
## Let's look at their GDP per Capita growth and TFP.
```{r}
library(readstata13)
pwt <- read.dta13("C:/Users/Jina Gil/Documents/SNU/Spring 2020/Econ. Growth/data/pwt91.dta")
pwt$gcap <- pwt$rgdpna / pwt$pop
pwt.g = pwt %>%
  arrange(countrycode) %>%
  mutate(dif.year = year - lag(year),
         dif.growth = gcap -lag(gcap),
         grate = (dif.growth / dif.year)/gcap*100)
argen <- pwt.g[pwt.g$countrycode %in% c("ARG"),]
chi <- pwt.g[pwt.g$countrycode %in% c("CHL"),]
brazil <- pwt.g[pwt.g$countrycode %in% c("BRA"),]
scl=20
```
---

## First, Argentina
```{r}
g_a <- ggplot(argen, aes(x=year, y=grate, color=countrycode))+
  geom_line()+theme_bw()+ geom_line(aes(y = ctfp*scl, colour = "TFP")) +
  scale_y_continuous(sec.axis = sec_axis(~./scl, name = "Index 2011=1"))+
  labs(title="GDP per Capita Growth Rate", subtitle="Argentina",
       x="Year", y="Growth (%)")
g_a + labs(colour = "factor")
```

## Second, Brazil 
```{r}
g_bra <- ggplot(brazil, aes(x=year, y=grate, color=countrycode))+
  geom_line()+theme_bw()+ geom_line(aes(y = ctfp*scl, colour = "TFP")) +
  scale_y_continuous(sec.axis = sec_axis(~./scl, name = "Index 2011=1"))+
  labs(title="GDP per Capita Growth Rate", subtitle="Brazil",
       x="Year", y="Growth (%)")
g_bra + labs(colour = "factor")
```

## Third, Chile
```{r}
g_c <- ggplot(chi, aes(x=year, y=grate, color=countrycode))+
  geom_line()+theme_bw()+ geom_line(aes(y = ctfp*scl, colour = "TFP")) +
  scale_y_continuous(sec.axis = sec_axis(~./scl, name = "Index 2011=1"))+
  labs(title="GDP per Capita Growth Rate", subtitle="Chile",
       x="Year", y="Growth (%)")
g_c + labs(colour = "factor")
```

## Wait, the results are...
- TFP, Growth rate goes almost together. Not surprising. 
(TFP? Time Factor Productivity, measuring the technology development contributing to GDP)
- Does this mean that there is a manufacturing sector, high-tech industry increasing TFP? We do not know. Yet. 

## What we get?
- Their main exports *are* commodity goods, yet they also exports manufactures!
- Their TFP level follows the GDP per Capita growth, so there might be...?
- Yet, we need to measure out the TFP **Growth** rate.
- Regression is needed! Yet, the dataset is not balanced, so need more variables to control it. 
